A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-04-22 papers/paper/preprint.tex
Edits to print out and work on

Critical infrastructure services in America are aging and inadequate, leading to high-impact failures even during conditions with substantial precedent in the historical record \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
A critical question, then, is what conditions or futures particular infrastructure systems and components should be designed to withstand.
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
However, this decision is also subject to assumptions regarding the likelihhod and relevance of possible futures.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, based on probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a nominal requirement for homes purchased through using federally backed mortgage porgrams to be covered by flood insurance \citep{kousky_voucher:2014}.
Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
Design standards area also used to design large-scale infrastructure.
For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is $<\nicefrac{1}{4000}$ years \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project proposed for  Galveston Bay, was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

Risk-based design \citep{eijgenraam_flooding:2014,vandantzig_dike:1956}\ldots\james{Maybe add a bit here -- this is a low hanging critique}
And homeowners subject high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.
Like standards-based approaches, cost-benefit analyses or risk-based designs rely on probabilistic descriptions of relevant hazard.

Current methods for estimating the probability distributions that underpin cost-benefit analysis or design events focus on data-driven, objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies the procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity duration and frequency of extreme rainfall.
A key assumption underlying these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
Yet changes in both global climate and local environmental conditions challenge this assumption in many cases \citep{Merz:2014gf,Milly:2008dg}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
While methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of the trend which may not represent the physical mechanisms that lead to nonstationarity or their uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.

@@

Many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
For example, robust decision methods \citep{lempert_shaping:2003} run models over large ensembles of possible futures to assess the performance of different policies, then apply statistical analysis to identify the conditions under which particular policies perform well or poorly.
These methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
When many possible levers are available, and exhaustive sampling therefore infeasible, multiobjective robust decision methods add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
In addition, many studies formally quantify robustness to deep unccertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.

Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses integrate performance over many possible futures, whether to compute sensitivity of responses to inputs, estimate expectations of functions, or compute robustness metrics, they necessarily make assumptions about likelihood of different scenarios.
Although many analyses contend that specifying a joint probability distribution function over inputs is inappropriate under deep uncertainty, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.

@@

Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a family of candidate models, and we find many parallels to deep uncertainty.
The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, fr which probability distributions offer a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a discussion of methods for model selection).
From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}

In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model choice.
We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
For example, this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}.
Further, that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''

We proceed as follows.
In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-explore}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
In \cref{sec:case-study} we describe the case study in detail.
Next in \cref{sec:results} we discuss results for each of the three theoretical lenses.
Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.

@@

Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
While we assume for simplicity that the decision space is are known and finite, this approach could be coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling avoids making explicit assumption about the likelihood different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
\klaus{Does this address your comment about linking back to observations?}

@@

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric''  requires estimating the expectation of some function $\pi$, which has the form
\begin{equation*}
    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
\end{equation*}
Because expectations are defined relative to a probability distribution, \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}.

In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
Since the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.

Many decision analytic tools used in climate risk management can be interpreted through this lens.
A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
For example, \ldots\james{Add a water resources and a flood example}
In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
As we shall discuss in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only using a scenario approach (\ie, choices of model structures or factors such as forcings we show as cases) without any explicit probability statement \citep[see][]{wong_nola:2017}.\james{Wording}

Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
\begin{equation}
    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
\end{equation}
If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.

Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.

@@

If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
Consequently, climate risk management often uses the outputs of structured simulations and analyses.
For example, many studies use outputs from \gls{cmip} simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.

@@

We present here the case where the $\psi_j$ are one-dimensional.
We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
    \end{cases}
\end{equation}
This step is illustrated in \cref{fig:grid-sketch}.
For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
This approach is widely used in analysis of tropical cyclones: first an ensemble of cyclones is simulated, and then probabilistic weights are calculated for each based on attributes such as intensity, landfall location, and radius \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.

@@

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
However, reasoning using subjective probability allows\ldots\james{Finish sentence}

@@

Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.

@@

The system model (``relationships'' in \cref{fig:xlrm}) is comprised of two key pieces.
The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
The second converts model converts a time series of annual expected damages into lifetime expected damages.

@@

We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.\james{Expand a bit on other work and priors for SLR.}
Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.

>>>>

Critical infrastructure services in America are aging and inadequate, leading to high-impact failures under conditions known to be possible \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
A critical question, then, is what conditions or futures particular infrastructure systems and components should be designed to withstand.
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
However, this decision is also subject to assumptions regarding the likelihhod and relevance of possible futures.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, based on probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a nominal requirement for homes purchased with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require new construction in flood zones to be elevated some margin above a nominal \gls{bfe}.
Design standards are also used to design large-scale infrastructure.
For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.\klaus{I understand this to mean that they are the same for everyone, even if they don't necessarily lead to equitable outcomes -- is that how you read this?}
However, there are also limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
Like standards-based approaches, quantitative cost-benefit analyses rely on probabilistic descriptions of relevant hazard.

Currently used methods for estimating these \glspl{pdf} emphasize data-driven, objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
Among several statistical assumptions these analyses make that have been recently called into question \citep[see][for a review]{Merz:2014gf} is stationarity.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are apparent or anticipated \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
Although some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).

@@

Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
For example, robust decision making \citep{lempert_shaping:2003} runs models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
In addition, many studies formally quantify robustness to deep uncertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.

Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
Although some have argued\klaus{Can I get away w/o calling out a particular paper here?} that specifying a joint \gls{pdf} over inputs is inappropriate under deeply uncertain conditions, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.

@@

Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a candidate space.
The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders, decisionmakers, and analysts \citep{gelman_philosophy:2013}.
Within this philosophical framework we find many parallels to the motivations and philosophies of \gls{dmdu}.

In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''

We proceed as follows.
In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
In \cref{sec:case-study} we describe the case study in detail.
Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.

@@

Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
Similarly, we assume that a fixed ensemble of \glspl{sow} are available for analysis, but in some cases additional simulations may add value.

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling avoids making explicit assumption about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or mapping a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
\klaus{Does this address your comment about linking back to observations?}

@@

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires estimating the expectation of some function $\pi$, which has the form
\begin{equation}\label{eq:expectation-is-probability}
    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
\end{equation}
In other words, \cref{eq:expectation-is-probability} shows that \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}, $p(s)$.

In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing Monte Carlo estimation of \cref{eq:expectation-is-probability} \citep{betancourt:2018}.

This framework can be applied to understand many existing analyses.
For example, if we define $M$ to be a probabilistic scenario (\eg, a choice of physical model and of an \gls{rcp} scenario, as we shall see in the next section), then we can rigorously quantify uncertainty \emph{within models}, but can only characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$.

Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
For example, if this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.

@@

If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.

@@

We present here the case where the $\psi_j$ are one-dimensional.
We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
    \end{cases}
\end{equation}
This step is illustrated in \cref{fig:grid-sketch}.
For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
This weighting approach parallels the joint probability method used in analysis of tropical cyclones \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.

@@

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis, even when the \glspl{sow} deliberately over- or under-sample some regions of the parameter space.
As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.

@@

Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.

@@

The system model (``relationships'' in \cref{fig:xlrm}) is comprised of two key pieces.
The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
The second model converts a time series of annual expected damages into lifetime expected damages.

@@

We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.

We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
\Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
